---
title: "Let's correlate the s**** out of the data"
author: "Riccardo Fusaroli"
date: "May 26, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Language in Depression - Part 2

In this exercise, based on real data, we will explore whether depression severity is associated with a specific way of speaking.

Background: Previous research conducted in the USA has shown that people with depression tend to speak differently from comparison groups (see Pennebaker, The secret life of pronouns). In particular, they tend to use fewer words, fewer positive words, more negative words and more past tense. A Danish student has decided to replicate these findings in the Danish population. The data you have come from his research.

### Structure of the exercise

- Loading the libraries
- Setting the directory and loading the data

Note that we know the dataset already, so we are skipping that part.

- Let's go after the hypotheses!
  - Word count and depression severity (plot + statistical summary)
  - Use of positive words and depression severity (plot + statistical summary)
  - Use of negative words and depression severity (plot + statistical summary)
  - Use of past tense and depression severity (plot + statistical summary)

### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with

```{r}

```

### Define your working directory and load the data

- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data: "Day2Exercise1.csv" (use read.csv(fileName))

```{r}

getwd()
locpath=
setwd(locpath)
Data =

```

### Make sure the variables are in the right format (e.g. numeric vs. factor)

In particular, pay attention to:
- participant id
- gender
- diagnosis

```{r}


```

### Exercise 1) Let's check the new variable: depression severity

When you look at depression severity, is it normally distributed? Discuss whether the full dataset can be analyzed or whether you should only select a subset of the participants.

```{r}

# Plot in the full sample

# Aggregate by diagnosis

# Subset the data (if relevant)

```

## Exercise 2

Write a short paragraph reporting the results + plots per each of the dependent variables. 

### Depression severity and Total words used

Let's try now to look at relations of data

So for example: do people with more severe depression use less words than less depressed people?

#### Plotting

Let's use a plot to check that out:
- create a plot
- define the relevant variables
- define the kind of plot
- add axes and text

What can you infer from the graph?

```{r}


```

#### Statistical testing

Is there a relation? How big is it? How likely is to see that relation in the data if the null hypothesis were true?

```{r}

#N.b. Make sure you insert the subset of the data if you think that is relevant

```

### Depression severity and % of positive words

Let's focus on more specific word usage: The % of positive words 

#### Plotting

Let's use a plot to check that out:
- create a plot
- define the relevant variables
- define the kind of plot
- add axes and text

What can you infer from the graph?

```{r}

```

#### Statistical testing

There seems to be a difference. But how big is it? And how likely is to see that difference if the null hypothesis were true?

```{r}
```

### Depression severity and % of negative words

Let's focus on more specific word usage: The % of negative words 

#### Plotting

What can you infer from the graph?

```{r}

```

#### Statistical testing

There seems to be a difference. But how big is it? And how likely is to see that difference if the null hypothesis were true?

```{r}

```

### Depression severity and % of past tense verbs

Let's focus on more specific word usage: The % of past tense verbs

#### Plotting

What can you infer from the graph?

```{r}

```

#### Statistical testing

There seems to be a difference. But how big is it? And how likely is to see that difference if the null hypothesis were true?

```{r}

```

Exercise 3) Go to http://guessthecorrelation.com/ and play a game. How much did you score?

Exercise 4) Go to http://www.tylervigen.com/spurious-correlations and have a bit of fun.

Optional Exercise 5) Explore other linguistic indexes

Optional Exercise 6) Add gender and age to the model
