---
title: "Let's correlate the s**** out of the data"
author: "Riccardo Fusaroli"
date: "May 26, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Language in Depression - Part 2

In this exercise, based on real data, we will explore whether depression severity is associated with a specific way of speaking.

### Structure of the exercise

- Loading the libraries
- Setting the directory and loading the data

Note that we know the dataset already, so we are skipping that part.

- Visualize the data:
  - Histograms (distribution)
  - Scatterplots
  - Barplot
  - Boxplot
  - Violinplot

- Quantitatively describe the data
  - Mean
  - SD
  - Median
  - Quartiles
  
- Let's go after the hypotheses!
  - Word count and depression
  - Use of positive words and depression (plot + statistical summary)
  - Use of negative words and depression (plot + statistical summary)
  - Use of past tense and depression (plot + statistical summary)

### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with

```{r}

```

### Define your working directory and load the data

- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data: "Day2Exercise1.csv" (use read.csv(fileName))

```{r}

getwd()
locpath=
setwd(locpath)
Data =

```

### Make sure the variables are in the right format (e.g. numeric vs. factor)

In particular, pay attention to:
- participant id
- gender
- diagnosis

```{r}


```

### Let's check the new variable: depression severity

Plot it and describe it by diagnosis. Is it normally distributed?

```{r}

# Plot in the full sample

# Aggregate by diagnosis

# Subset the data (if relevant)

```

### Depression severity and Total words used

Let's try now to look at relations of data

So for example: do people with more severe depression use less words than less depressed people?

#### Plotting

Let's use a plot to check that out:
- create a plot
- define the relevant variables
- define the kind of plot
- add axes and text

What can you infer from the graph?

```{r}


```

#### Statistical testing

Is there a relation? How big is it? How likely is to see that relation in the data if the null hypothesis were true?

```{r}

# N.b. Make sure you insert the subset of the data if you think that is relevant

```

### Depression severity and % of positive words

Let's focus on more specific word usage: The % of positive words 

#### Plotting

Let's use a plot to check that out:
- create a plot
- define the relevant variables
- define the kind of plot
- add axes and text

What can you infer from the graph?

```{r}

```

#### Statistical testing

There seems to be a difference. But how big is it? And how likely is to see that difference if the null hypothesis were true?

```{r}
```

### Depression severity and % of negative words

Let's focus on more specific word usage: The % of negative words 

#### Plotting

What can you infer from the graph?

```{r}

```

#### Statistical testing

There seems to be a difference. But how big is it? And how likely is to see that difference if the null hypothesis were true?

```{r}

```

### Depression severity and % of past tense verbs

Let's focus on more specific word usage: The % of past tense verbs

#### Plotting

What can you infer from the graph?

```{r}

```

#### Statistical testing

There seems to be a difference. But how big is it? And how likely is to see that difference if the null hypothesis were true?

```{r}

```