---
title: "Exercise on generalized mixed effects models"
author: "Riccardo Fusaroli"
date: "May 27, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Language development in Autism Spectrum Disorder (ASD)

In this exercise we will explore language development in US children with and without ASD.

### Structure of the exercise

Basic stuff:
- Loading the libraries
- Setting the directory and loading the data
- Look at the data (which variables are there? Are they in the right format)
- Describe the participants (by diagnosis)

We will try to answer two separate questions:

- Do children with ASD develop language differently from non-ASD children?
- Do parents speak differently to children with ASD than to non-ASD ones?
  
### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with
```{r}

```

### Define your working directory and load the data

- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data: "Day2Exercise1.csv" (use read.csv(fileName))

```{r}

getwd()
#locpath=
setwd(locpath)
#Data =

```

### Visualize the dataset 

Identify relevant variables: participants demographic characteristics, diagnosis (ASD, ADOS, EL, Mullen) Visit, Number of unique words used, length of utterance in both child and parents.

Make sure the variables are in the right format

```{r}

```

### Characterize the participants

Remember to check them by diagnosis

```{r}

```

## Let's test hypotheses

### Hypothesis 1a: Unique words used by the child change: i) over time, ii) according to diagnosis

Remember to plot the data first and then to run a statistical test.

```{r}

```

### Hypothesis 1b: Length of utterance used by the child change: i) over time, ii) according to diagnosis

```{r}

```

### Hypothesis 2a: Unique words used by the child change: i) over time, ii) according to diagnosis

Remember to plot the data first and then to run a statistical test.

```{r}

```

### Hypothesis 2b: Length of utterance used by the child change: i) over time, ii) according to diagnosis

```{r}

```