---
title: "T-testing the s*** out of your data"
author: "Riccardo Fusaroli"
date: "May 23, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Language in Depression

In this exercise we will explore whether depression is associated with a specific way of speaking.

### Structure of the exercise

- Loading the libraries
- Setting the directory and loading the data
- Look at the data (which variables are there?)

- What to look for:
  - External validity (characteristics of the participants)
  - Hypothesis driven exploration (plot the relevant variables)

- Visualize the data:
  - Barplot
  - Boxplot
  - Violinplot

- Quantitatively describe the data
  - Mean
  - SD
  - Median
  - Quartiles
  
- Let's go after the hypotheses!
  - Word count and depression
  - Use of positive words and depression (plot + statistical summary)
  - Use of negative words and depression (plot + statistical summary)
  - Use of past tense and depression (plot + statistical summary)

### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with

```{r}

```

### Define your working directory and load the data

- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data: "Day2Exercise1.csv" (use read.csv(fileName))

```{r}

getwd()
#locpath=
setwd(locpath)
#Data =

```

### Visualize the dataset

Take a look at the data
- look at the head (headers plus first 6 lines)
- get number of rows, number of columns
- look at one row, one column
- look at one category (f.ex. diagnosis)
- select all participant with a diagnosis

```{r}

# Take a look at the first 6 lines
head(Data)
# How many datapoints?
nrow(Data)
# How many features?
ncol(Data)

# Select row 3 only
Data[3,]
# Select column 4 only
Data[,4]

# Visualize a specific feature
Data$diagnosis

# Select only participants with depression
Data[Data$diagnosis==1,]
```

### Make sure the variables are in the right format (e.g. numeric vs. factor)

In particular, pay attention to:
- participant id
- gender
- diagnosis

```{r}


```

### Quantify the dataset

You can also get a quantitative overview over the different variables (Min, 1st Quartile, Median, Mean, 3rd Quartile, Max) using summary() and sd()

```{r}

```

### Let's get to business - checking the participant sample

It's important to get a feel for your data and plots are often the best way to do so.

First we want to check a few things:

- How many participants do we have? How many with a diagnosis? How many in the comparison group?
- What about age and gender? Are they equally represented in the data? and in the two groups?

#### We start with number of participants and by group

This is simple. Let's just summarize the column diagnosis

```{r}

cdata <- aggregate(Data["nr"], by=Data[c("diagnosis")], FUN=length)

```

#### What about gender? In the whole sample and in each group


```{r}


```

#### What about age? In the full sample and by diagnosis?

- Create the plot and specify the variables at stake
- Define which kind of plot
- Add text to the plot
- Calculate it (using aggregate)

What can we infer from the graph? And from the calculation?

```{r}

# Boxplot

plot2 <- ggplot(Data) + geom_boxplot(aes(x=diagnosis,y=age, fill=diagnosis)) +
          ggtitle("Boxplot representing the participants' age by diagnosis") + 
          xlab("Diagnosis") + 
          ylab("number of participants") + 
          scale_x_discrete(labels=c("Comparison", "Depression"))
print(plot2)

# Violin Plot

# Use aggregate to calculate mean and sd of age by diagnosis

```  

#### What about education? In the full sample and by diagnosis?

What can we infer from the graph? And from the calculation?

```{r}

```


### From participants to relations: Total words used

Let's try now to look at relations of data

So for example: do people with depression use less words than not depressed people?

#### Plotting

Let's use a plot to check that out:
- create a plot
- define the relevant variables
- define the kind of plot
- add axes and text

What can you infer from the graph?

```{r}


```

#### Statistical testing

There seems to be a difference. But how big is it? And how likely is to see that difference if the null hypothesis were true?

```{r}
model1=lm(totalWords~diagnosis,Data)
summary(model1)
```

### From participants to relations: % of positive words

Let's focus on more specific word usage: The % of positive words 

#### Plotting

Let's use a plot to check that out:
- create a plot
- define the relevant variables
- define the kind of plot
- add axes and text

What can you infer from the graph?

```{r}

```

#### Statistical testing

There seems to be a difference. But how big is it? And how likely is to see that difference if the null hypothesis were true?

```{r}
```

### From participants to relations: % of negative words

Let's focus on more specific word usage: The % of negative words 

#### Plotting

What can you infer from the graph?

```{r}

```

#### Statistical testing

There seems to be a difference. But how big is it? And how likely is to see that difference if the null hypothesis were true?

```{r}

```

### From participants to relations: % of past tense verbs

Let's focus on more specific word usage: The % of past tense verbs

#### Plotting

What can you infer from the graph?

```{r}

```

#### Statistical testing

There seems to be a difference. But how big is it? And how likely is to see that difference if the null hypothesis were true?

```{r}

```

### What's next?

Is depression a yes/no diagnosis? What about severity of clinical features?  [Tomorrow]

  - Word count and depression severity
  - Use of positive and depression severity (plot)
  - Use of negative and depression severity (plot)
  - Use of past tense and depression severity (plot)

